---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rockup

<!-- badges: start -->
[![GitHub issues](https://img.shields.io/github/issues/rytenlab/rockup)](https://github.com/rytenlab/rockup/issues)
[![GitHub pulls](https://img.shields.io/github/issues-pr/rytenlab/rockup)](https://github.com/rytenlab/rockup/pulls)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check-bioc](https://github.com/rytenlab/rockup/workflows/R-CMD-check-bioc/badge.svg)](https://github.com/rytenlab/rockup/actions)
[![Codecov test coverage](https://codecov.io/gh/rytenlab/rockup/branch/main/graph/badge.svg)](https://app.codecov.io/gh/rytenlab/rockup?branch=main)
<!-- badges: end -->

The goal of `rockup` is to make it easy to connect to R Studio Server via [rocker](https://github.com/rocker-org/rocker).

## Installation instructions

Install the development version of `rockup` from [GitHub](https://github.com/rytenlab/rockup) with:

```{r 'install_dev', eval = FALSE}
BiocManager::install("rytenlab/rockup")
```

## Basic usage

`rockup` has a single user-level function [docker_run_rserver](https://rytenlab.github.io/rockup/reference/docker_run_rserver.html), which is designed to start a container running R Studio Server on a [rocker](https://github.com/rocker-org/rocker). Below is an example of how to run `docker_run_rserver` with minimal config, for a more detailed guide see the [vignette](https://rytenlab.github.io/rockup/articles/rockup.html). 

```{r basic-usage, eval = FALSE}
docker_run_rserver(
  image = "bioconductor/bioconductor_docker:RELEASE_3_13",
  port = 8888,
  name = "your_container_name"
)
```

## Code of Conduct

Please note that the `rockup` project is released with a [Contributor Code of Conduct](http://bioconductor.org/about/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

## Development tools

* Continuous code testing is possible thanks to [GitHub actions](https://www.tidyverse.org/blog/2020/04/usethis-1-6-0/)  through `r BiocStyle::CRANpkg('usethis')`, `r BiocStyle::CRANpkg('remotes')`, and `r BiocStyle::CRANpkg('rcmdcheck')` customized to use [Bioconductor's docker containers](https://www.bioconductor.org/help/docker/) and `r BiocStyle::Biocpkg('BiocCheck')`.
* Code coverage assessment is possible thanks to [codecov](https://codecov.io/gh) and `r BiocStyle::CRANpkg('covr')`.
* The [documentation website](http://rytenlab.github.io/rockup) is automatically updated thanks to `r BiocStyle::CRANpkg('pkgdown')`.
* The code is styled automatically thanks to `r BiocStyle::CRANpkg('styler')`.
* The documentation is formatted thanks to `r BiocStyle::CRANpkg('devtools')` and `r BiocStyle::CRANpkg('roxygen2')`.

For more details, check the `dev` directory.

This package was developed using `r BiocStyle::Biocpkg('biocthis')`.
