---
title: "Connecting to R Studio server using rockup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rockup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rockup)
```

<br><br>

# Background

- [R Studio server](https://www.rstudio.com/products/rstudio/download-server/) allows user's to code using a convenient IDE, whilst harnessing the processing power of a computing remote cluster/server. 
- However, the installation and maintenance of the R Studio server software can be time-consuming. Also, the free version of R Studio server only permits a single version of `R` to be used. Assuming you are not willing to pay, this can make analyses that depend on `R` packages across multiple versions of `R` inconvenient to run. 
- Using a docker image from the [rocker project](https://hub.docker.com/r/rocker/rstudio) with `R`/R Studio server pre-installed, you can easily bypass the two limitations above. A user can connect to the `rocker` image via an `ssh` tunnel, accessing R Studio server through their local web browser. Each docker image can have a unique installation of R Studio server, thus any number of `R` versions can be used. 
- The following guide will demonstrate how to install and connect to a R Studio server container that's running on a `rocker` image on a remote server. 
- This guide does **not** cover the fundamentals of docker itself and it is recommended that anyone using this guide should already have a basic proficiency with docker.  

<br><br>

# Guide

## Docker installation

You must have `docker` installed on your system. To check you have `docker` installed, you can use: 

```{bash check-docker-version, eval = FALSE}
# based on: https://www.digitalocean.com/community/questions/how-to-check-for-docker-installation
docker -v
echo $?
```

If you don't, install `docker`. A guide to installing `docker` on Ubuntu 18.04 can be found [here](https://www.digitalocean.com/community/tutorials/how-to-install-and-use-docker-on-ubuntu-18-04).

<br><br>

## Download a `rocker` image

In order to use R Studio server, a `docker` image with R Studio server pre-installed is required. [Bioconductor](https://www.bioconductor.org/help/docker/) releases it's own `docker` image that inherits from `rocker`, with other useful system dependencies for analyses of biological data pre-installed, such as core Bioconductor packages. You can download the Bioconductor docker image using:

```{bash download-rocker, eval = FALSE}
# the 3.13 release version of Bioconductor docker image is used here
# be sure to check for an updated version as and when you use this guide
sudo docker pull bioconductor/bioconductor_docker:RELEASE_3_13
```

<br><br>

## Start a R Studio server process (on a remote server)

To initialise a container running R Studio server on the Bioconductor `docker` image downloaded above, you need to execute the `docker run` command with various flags depending on your desired setup. `rockup` has a single user-level function `docker_run_rserver()` which is designed to make this step easier and executable within `R`. 

First, make sure you have `rockup` installed. If you don't, you can install `rockup` from GitHub via:

```{r install-rockup, eval = FALSE}
# this requires R version >= 4.0
devtools::install_github("ryten_lab/rockup")
```

Next, run `rockup::docker_run_rserver()` within an `R` terminal **on the remote server/machine where you want to use R Studio server**. At a minimum, you should set the `image`, `port` and `name` arguments explained below. Setting the `verbose` argument to TRUE will print the flags that were used within the `docker` command and can be useful for debugging or logging. 

**Important**: if you are running this on the Ryten lab server, it is worth making sure: 1. The `port` you use does not overlap with one that's already in use. The list of in use ports can be found in the ryten_induction bookdown [here](https://rytenlab.github.io/ryten_induction/server-usage.html#ryten-ports). 2. If you're occupying a new `port`, please add it to the list of in use ports (or ask another member to).

```{r docker-run-rserver-ex1, eval = FALSE}
rockup::docker_run_rserver(
  image = "bioconductor/bioconductor_docker:RELEASE_3_13", # a rocker image, or one that inherits from rocker
  port = 8888, # port on which the host will have present R Studio server
  name = "example", # name of docker process
  verbose = TRUE # whether to print out the flags passed to the docker command
)
```

<br><br>

## Connecting to a R Studio server process (from your local machine)

Once the R Studio server process is running, the next step is to map the `localhost` port of your local machine to the `port` presenting R Studio server on the remote server (specified above as 8787). You can do so by executing the `ssh` command shown below **on your local machine**: 

```{bash ssh-tunnel, eval = FALSE}
# make sure the port specified here matches the port you have used above in 
# rockup::docker_run_rserver()
ssh \
-X -N -f -L localhost:8787:localhost:8787 \
user@ip
```

If the above `ssh` command has run successfully, you will now be able to access R Studio server by going to the address `localhost:8787` on your local browser. The default login details for the Bioconductor docker are:

Username: **rstudio**

Password: **bioc**

More details of the Bioconductor docker can be found [here](https://www.bioconductor.org/help/docker/).

<br><br>

## Mounting volumes 

Most analyses relies on data that is stored on the original host machine, therefore not (by default) accessible by the `docker` container. Therefore, it is often useful to mount the required files, allowing them to be accessible by the `docker` process. Mounting files can be done using `rockup::docker_run_rserver()` via the arguments `volumes`, `volumes_ro`. 

Directories mounted using the `volumes_ro` argument will always have read-only access permissions on the `docker` container. 

The user permissions for accessing files mounted using `volumes` argument are dictated by the `USERID` and `GROUPID` arguments. These should be set matching the user you would like to mirror the permissions of. On Linux, the `USERID` and `GROUPID` of the current user can be obtained via the `bash` command `id`. **Warning: depending on the permissions, this could give your docker container the ability to delete the mounted files**

Below is an example of running `rockup::docker_run_rserver()`, whilst mounting volumes: 

```{r docker-run-rserver-ex2, eval = FALSE}
# volumes - paths will be mounted with user permissions
# matching user specified by the USERID and GROUPID arguments
# volumes_ro - paths will be mounted with read-only access
rockup::docker_run_rserver(
  image = "bioconductor/bioconductor_docker:RELEASE_3_13",
  port = 8888,
  name = "example_2",
  verbose = TRUE,
  volumes = c(
    "/path/to/mounted/dir/that/will/match/permissions/"
  ),
  volumes_ro = c(
    "/path/to/mounted/dir/that/will/be/read_only"
  ),
  permissions = "match",
  USERID = 1000,
  GROUPID = 1000
)
```

<br><br>

# Session info

<details>
  <summary>Show/hide</summary>
```{r session-info, echo = FALSE}

library("sessioninfo")

options(width = 120)

session_info()
```
</details> 
